---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Te

<!-- badges: start -->
<!-- badges: end -->

The goal of Te is to perform "exact" two sample test with unequal variance.

$$X_i \sim N(\mu_1,\sigma_1^2), i=1,2,...,n_1$$

$$Y_i \sim N(\mu_2,\sigma_2^2), i=1,2,...,n_2$$

Hypothesis testing when $\sigma_1^2 \not = \sigma_2^2$:

$$H_0: \mu_1 = \mu_2$$

# History
Welch test is frequently used for two sample t test, when two groups' variances are unequal. However it is an asymptotic method, which means the type one error is not exactly alpha. 

Te test (e for exact) performs exact/non-asymptotic two sample t test, where two groups' variance can be unequal and the type one error is exactly alpha. It achieves the maximal degree of freedom and has better power than simple paired t test.

Go to https://arxiv.org/pdf/2210.16473.pdf for more details.

## Installation

You can install the development version of Te like so:

```{r}
library(devtools)
install_github("hobbitish1028/Te_test",force=TRUE)
library(Te)
```


## Example

This is a basic example which shows you how to solve a common problem:

```{r}
library(Te)
n1<-5
n2<-50
mu1<-0
mu2<-0
sigma1<-1
sigma2<-2
x<-rnorm(n1,mu1,sigma1)
y<-rnorm(n2,mu2,sigma2)
Te_test(x,y)
```

Find confidence interval for $\mu_1 - \mu_2$

```{r}
Te_CI(x,y,alpha=0.05)
```


## Simulation

Here we also show a quick simulation to compare the performance between Te test, Welch test and Z test. We can see Welch's type one error is biased from 0.05, which means it is not an exact test. Te has a type one error close to alpha = 0.05 bacause it is an exact test.

Although Welch t test is a robust one, we can see that Te test outperforms it significantly when sample sizes are relative small (<100) and two size differs a lot (e.g. 5 and 50).

```{r}
set.seed(123)
res<-sim_Te(mu1=0,mu2=0,sigma1=1,sigma2=2,n1=5,n2=50,rpt=1e5)
res$type_one_error
```
